### 2023-01-04 Lab1_R_Review_Answers.Rmd
Add files via upload

---
title: '784 Lab 1: R Refresher: Answer Key'
author: "Eric Parajon"
output:
  pdf_document:
    keep_tex: TRUE
  html_document:
    df_print: paged
---

Code from today's lab was partially adapted from labs by Simon Hoellerbauer and Isabel Laterzo.

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = TRUE, include = TRUE)
```

Today's lab will help you refresh some of what you learned in R in 783. It will also introduce you to some topics in R that you may not be familiar with, but are very useful for this semester.

For a refresher on how to use R Markdown, please see this cheat sheet: https://rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf.

Within your code chunks, comment our your code describing what you are doing in each step. This is an important habit to establish.

#R Review
\medskip
1. First, it is important to save the document you are using in a central folder. Then, set your working directory to that folder. It's also important to clear your environment before working in a new document.
```{r, include = TRUE, eval = FALSE}
getwd() #what is your current working directory?

setwd() #in the parentheses, set the working directory to the folder where
#you save this file. Realistically, this isn't super relevant now because
#you're not reading in external data, but it's a good habit to get into!

rm(list=ls()) #clear global environment
```

\bigskip
2. First, create a vector of ten 5s using the function `rep()`. Second, create a vector that counts from 1 to 10 - do this in two ways, one using `c()` and one using `seq()`. Third, create a vector of length 10 filled with NAs. Finally, uses `cbind()` to bind theses four vectors together to create a 10 x 14 (10 rows, 4 columns) matrix. Remember, if you don't know how to use a function, type ? function name into your console to see the help file (e.g., `?seq()`).

Try binding the vectors together with `rbind()` instead. What happens?

```{r}

vec1 <- rep(5, 10) #vector of ten 5s

vec2 <- c(1:10) #vector counting from 1 to 10

vec3 <- seq(from = 1, to = 10, by = 1) #vector counting from 1 to 10

vec4 <- rep(NA, 10) #vector of 10 NAss

all_vecs <- cbind(vec1, vec2, vec3, vec4) #bind vectorss together using cbind
#we get four columns and ten rows

all_vecs2 <- rbind(vec1, vec2, vec3, vec4) #bind vectors together using rbind
#we get ten columns and four rows

```

\bigskip
3. Set your seed to 425. Simulate 100 draws from a Normal distribution with mean 2 and standard deviation 4, assigning the result to an object.

What proportion of the draws is above 3.1 in our sample? Think about using the functions `sum()` and `length()` to solve this, although there are other ways as well.

What is the expected proportion of draws above 3.1? Hint, look into the function `pnorm()`

Repeat this process, but this time take 10,000 draws from the same Normal distribution. How does our result compare with the expected proportion of draws above 3.1 now?

Thinking back to last semester, what did we just calculate using simulation?

```{r}
set.seed(425)

#simulate 100 draws from N(2, 4)
draws <- rnorm(100, 2, 4)

#calculate proportion of draws above 3.1
sum(draws > 3.1)/length(draws)
#0.39

#using pnorm to retrieve the expected proportion of draws above 1.5
1 - pnorm(3.1, 2, 4)
#just about matches above!

#repeating the above with 100,000 draws
draws2 <- rnorm(10000, 2, 4)
sum(draws2 > 3.1)/length(draws2)
#0.3864, very similar

#We simulated 1 - CDF of N(2, 4) at x = 3.1

```

\bigskip
4. Another important concept in R is a `for` loop. If you need a refresher on these, visit: https://www.r-bloggers.com/2015/12/how-to-write-the-first-for-loop-in-r/.

Using the 100 draws from the Normal distribution from Q3, create a new vector that is 1 when the vector of draws is greater than 3.1 and 0 when it is less than or equal to 3.1. First, do this using a `for` loop with an embedded `if` `else` sequence. What is the sum of this vector?

In addition to loops, `ifelse()` is also an important function. Look at the documentation of this function and repeat the above activity using `ifelse()`.

```{r}
#creating a container (here, a vector) into which we can dump the 1s and 0s. 
#We could do this with c() but it's more efficient to create
#a container of the proper size first 
draws_new <- vector("integer", length(draws))

#for loop that fills the container with 1 when draws > 3.1 and 0 elsewise
for (i in 1:length(draws)){
  if(draws[i] > 3.1){
    draws_new[i] <- 1
  } else draws_new[i] <- 0
}

sum(draws_new) #the number here looks familiar, no?

#repeat above using ifelse
draws_new2 <- ifelse(draws > 3.1, 1, 0)

#checking to make sure we have the same number
sum(draws_new2)
#yay! it is the same.

```

\bigskip
5. Load in the `swiss` dataset. A description for this data can be found here: https://stat.ethz.ch/R-manual/R-patched/library/datasets/html/swiss.html How many variables does it have? How many observations?

The variable `Catholic` is the \% of Catholic individuals, as opposed to Protestant in each province. First look at the summary of this variable. Then, create a new variable `Protestant` that is the percent of Protestant individuals in each province, just going with the assumption that all non-Catholics are Protestant here.

What is the mean percentage of this new variable?

```{r}
#loading in swiss dataset
data(swiss)

#find number of variables and number of observations
ncol(swiss); nrow(swiss)
#six variables, 47 observations

#summary
summary(swiss$Catholic)

#creating protestant variable
swiss$Protestant <- 100 - swiss$Catholic

#summary
summary(swiss$Protestant)

#finding mean of Protestant
mean(swiss$Protestant)
#58.86

```
\bigskip
6. Now, find out what type of object `swiss` is. Then, coerce `swiss` to a list by using `as.list()` and assign it to an objected named `swiss_list`. Look at the object with `View()` What happened? Use the `length()` function on `swiss`. What is the output? Why?

Create a matrix filled with random values that has 4 rows and 2 columns. Use `length()` on this matrix. Why might this be? Use `?` to read the documentation for `length()`.
```{r}
#finding the type of swisss
class(swiss) 
typeof(swiss)

#corece swiss to a list
swiss_list <- as.list(swiss)

#View
View(swiss_list)

#get length of swiss dataset 
length(swiss)

#7 - this is because swiss is a dataframe and a dataframe is really a list of
#same length vectors. This fact helps dataframes mimic matrices.

#this is an example matrix
test_mat <- matrix(c(1, 0, 3, 2, 4, 5, 6, 8), nrow = 4, ncol = 2)
#this is the length
length(test_mat)
#8

#?length()
```
\bigskip
7. Continuing to use the original `swiss` data, use a for loop to calculate the median of each of the variables. Store the result in a *named* vector (i.e., provide a name to each of the values so they are not just a bunch of random numbers with no context!).

Some useful functions here are `vector()`, `seq_along()`, and `names()`. Remember to use the help function `?` for further information about these functions.
```{r}
#creating container for for loop
medians <- vector("double", ncol(swiss))

#for loop that finds the median for each variable (column)
for (i in seq_along(swiss)){
  medians[i] <- median(swiss[, i])
}

#assigning the names of swiss (variable names/column headers) to the median
#vector
names(medians) <- names(swiss)

#print
print(medians)
```
\bigskip
8. Look into the three functions `apply()`, `sapply()`, and `lapply()`. What are the differences between these functions?

Repeat 7, but this time use the three above functions. Using each of these functions, examine what type of object they produce. How are they different from using a `for` loop? How are they different from each other in terms of what they produce? A useful function for examining different object types is `class()`.
```{r}
#examine functions
?apply()
?sapply()
?lapply()

#using apply
medians2 <- apply(swiss, 2, median)
print(medians2)
typeof(medians2)
#it created a named vector

#using lapply
medians3 <- lapply(swiss, median)
print(medians3)
typeof(medians3)
#outputs a named list

#using sapply
medians4 <- sapply(swiss, median)
print(medians4)
typeof(medians4)
#outputs a named vector
```
\bigskip
9. Now we will explore `ggplot2`, which is a great (but sometimess tedious) package for plotting data and results. You will probably grow to have a love/hate relationship with `ggplot2`. For help with this package, reference this cheat sheet: https://rstudio.com/wp-content/uploads/2016/11/ggplot2-cheatsheet-2.1.pdf

Using `ggplot`, create a scatterplot with the `swiss` dataset that plots `Education` against `Examination`. Add axes labels and a title. Does it look like there might be a relationship between these two variables?
```{r}
library(ggplot2)

ggplot(swiss, aes(x = Education, y = Examination)) +
  geom_point() + 
  labs(x = "Education", y = "Examination", title = "Scatterplot") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

#looks there there could be a positive relationship between Examination and Education 
# as Examination score increases, Education also increases

```

\newpage

\bigskip
10. It is also useful to be able to write your own functions in `R`. Sometimes, `R` doesn't have a function for exactly what you are looking for. Doing so is pretty straightforward. For a quick review, see:  https://swcarpentry.github.io/r-novice-inflammation/02-func-R/

For example, if I wanted to write a function that evaluated the equation \[f(x, y) = xy \] I could write one of the following options:

```{r, include = TRUE}
#example 1
fun_ex1 <- function(params) {
  x <- params[1]
  y <- params[2]

  result <- x*y

  return(result)
}

#evaluate
fun_ex1(c(3,4))

#example 2
fun_ex2 <- function(x, y) {

  result <- x*y

  return(result)
}

#evaluate
fun_ex2(3,4)

```

Now your turn. Write a function that evaluates the following equation:
\[f(x, y, z) = x^2 + 3xy + \sin(z)
\]
Have this function takes a vector as its single argument (see example 1), made up of the three parameters ($x, y, z$). Evaluate the function at $x = 3/4, y = 13, z = 3.4$.

```{r}
#function for the above equation
my_fun <- function(params){
  x <- params[1]
  y <- params[2]
  z <- params[3]

  result <- x^2 + 3*x*y + sin(z)

  return(result)
}

#evaluating the function at x = 3/4, y = 13, z = 3.4
my_fun(c(3/4, 13, 3.4))
```

\newpage
# Math Review:
\bigskip
11. Let's try out some linear algebra. Below are 2 vectors, a and b, and two matrices A and B. Use them for this section.
Here is a link to a linear algebra cheat sheet in R (and other languages): 
https://github.com/scalanlp/breeze/wiki/Linear-Algebra-Cheat-Sheet
```{r, include = TRUE}
#a
a <- c(4, -5, 4, 1)
#b
b <- c(3, 2, -7, 17)

#A
A <- matrix(c(1, -3, 4, 5, 6, 8, 9, -10), ncol = 4)

#B
B <- matrix(c(4:1, 8, 14, -2, 1), ncol = 2)
```

What is the dot product of $\mathbf{a}$ and $\mathbf{b}$? 
Does the result change if we take the dot product of $\mathbf{b}$ and $\mathbf{a}$?

What are the dimensions for the matrix $\mathbf{A}\mathbf{B}$? Is this new matrix invertible?

What are the eigenvectors and eigenvalues of this new matrix?

Verify that the values produced are actually eigenvalues.

```{r}
#dot product
a %*% b
b %*% a

#AB
AB <- A %*% B
solve(AB)
dim(AB)
#yes it is invertible

#eigen vectors and values
eigen_AB <- eigen(AB)

#verifying that they are eigenvalues
sapply(eigen_AB$values, function(x) det(AB - diag(2)*x))

```