### 2023-01-13 Lab 3  Fitting LMs/Lab3_LM_Answers.Rmd

---
title: 'Lab 3: The Linear Model in R: Answer key'
author: "Eric Parajon"
output:
  html_document:
    df_print: paged
---

Adapted  from code and text from Isabel Laterzo, Simon Hoellerbauer and Chelsea Estancona.

This lab will introduce you to `lm()`, the function we will be using to fit linear models (for now) in \texttt{R}. I will walk you through the different parts of this lab, but then we will largely be working independently. Chime in whenever you have questions!

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = TRUE, include = TRUE, echo=TRUE, eval=TRUE)

#Clear workspace
rm(list=ls())

#The pacman package is required to load/install the additional packages. Uncomment the line below to install the package.
#install.packages("pacman")

#Installing and loading packages for use
pacman::p_load(xtable,stargazer,palmerpenguins,tidyverse,texreg)

```

Don't forget to set your working directory!
```{r, eval = FALSE}

setwd()

```

# Part 1: LM in R

+ First, we'll simulate some data to use. This will be a very useful skill in the future, so pay attention to what we do here.

+ We first create our variables and then stipulate what our effects should be.

+ Let's set our seed to 123.

+ For our $\mathbf{x}_1$ variable, let's take 20 draws from a $\mathcal{N}(1.3, 2.3)$ (using the `R` parameterization, where the second parameter is the standard deviation).

+ Let's set $\beta_0$ to be 17, and $\beta_1$ to be 1.3. We also can't forget to simulate $\boldsymbol{\epsilon}$! Let's $\boldsymbol{\epsilon} \sim \mathcal{N}(0, 3.4)$.

```{r}
set.seed(123)

##Explanatory variables
x1 <- rnorm(20, 1.3, 2.3)

##Coefficients
b0 <- 17 #Intercept
b1 <- 1.3 #Beta 1

##standard deviation
sigma <- 3.4

error <- rnorm(length(x1), 0, sigma) #Let's simulate some random error, too. 
#Random error is always included in a regression model because unless there is a linear relationship between the variables in the model, there will always be some amount of error (or variation) which is not totally unaccounted for by the model.
```

+ What would we do in order to simulate a linear relationship with the variables and effects above? Create the outcome variable `y`.
```{r}
y <- b0 + b1 * x1 + error
```

+ What do $\mathbf{x}_1$ and $\mathbf{y}$ look like together? Use the `plot` function to plot a scatter plot of $\mathbf{x}_1$ against $\mathbf{y}$.

```{r}
plot(x1, y)
```

+ Now that we have simulated explanatory variables and set up a 'true' relationship between our explanatory variable (x1) and outcome variable, we can use R's built-in functions to evaluate our linear model. For a linear model, we will want to use the `lm()` function. The basic syntax for this function is `lm(Y ~ X1 + X2)`. Use `summary()` on the model object (`m1` in this case) in order to see some of the most interesting information about the fitted model.

```{r}
m1 <- lm(y ~ x1) #the basic syntax here: lm(Y ~ X1 + X2)
#summarize the 'model object' we've created
summary(m1)
```

# Part 2: Combining LM and the tidyverse
+ We can also combine what we learned about using the tidyverse with linear modeling!

+ As we dive into learning linear models in `R`, we're going to use some data from the `mtcars`. But first, lets use the `tidyverse` to clean it up a bit. First, take a look at the mtcars data (its readily available so you don't really need to load it in!).

```{r}
library(tidyverse)
#examine data
summary(mtcars)
```

+ Great. Let's say we want to analyze the role that `disp` and `cyl` play in influencing `mpg`. However, we're only interested in observations where `cyl` is greater than 4. Let's use `dplyr` and the `tidyverse` to only keep observations where `cyl` is greater than 4. Then, only keep the three above variables in the new data set. Label this data set `new_data`.

+ Hint: the `dplyr` functions you are interested in here are `filter()` and `select()` - don't forget to call their help files if you need more information (e.g., `?filter()` in your console).

```{r}

new_data <- mtcars %>%
  filter(cyl > 4) %>%
  select(cyl, disp, mpg)
```

+ Now that we have our data ready, we can use `lm()`.  Fill in the blank in the chunk below where Y is `mpg`, X1 is `cyl`, and X2 is `disp`. Use `summary()` on the model object (`m2` in this case) in order to see some of the most interesting information about the fitted model.

+ Remember, our data is called `new_data`.

```{r}
#model
m2 <- lm(mpg~disp+cyl, data=new_data)

#summarize
summary(m2)

```

+ The `summary` function provides lots of valuable information about the model we've created. We can also extract different information if we need, for example, just the coefficients, or just the residuals. Use `View()` on `m2`. What does it look like `m2` is? What do you think we could do if we wanted to access the different objects within `m2`?

```{r, eval=FALSE, echo=TRUE}
#Taking a look at m2
View(m2)

#extract coefficients
m2$coefficients

#extract residuals
m2$residuals

#extract y hat 
m2$fitted.values
```

+ We can also use indexing to get to parts of these vectors - the output from a model is a list object, which means we can use brackets with $ indexing.

```{r}
m2$coefficients[1] #What does this give us?
m2[2] #What about this?
```

+ What if we wanted to include an interaction term?
```{r}
m3<-lm(mpg~disp+cyl+disp*cyl, data=new_data) #use * for interaction- include both variables

summary(m3)
```

+ We don't actually have to write all of the variables two times; the `lm` function will include the variables involved in the interaction even if we don't specify them individually in the formula. Try this out below, creating model object `m4` and compare it to `m3`.
```{r}
#m3:
m4 <- lm(mpg ~ disp*cyl, data=new_data)

summary(m4)
```

+ How might we present these results in a table? Both `xtable()`, `stargazer()`, and `texreg()` are packages that produce LaTeX code that could be copy-pasted into a .tex document or included in an Rmarkdown document. Try these on one of the model object `m4` we created above and see what happens. Note, these are VERY flexible and can be customized to make very nice tables, the ones we are generating here are just basic. 
```{r, results='asis'}
#These produce code for a .pdf that we can copy and paste
xtable(m4)
stargazer(m4,type="html")
htmlreg(m4)

#you can also add lots of useful information onto these
#check out guides online!

```

+ When presenting our results, we might be largely concerned with prediction: if I have a model and provide it with 'new' data, what should I expect? How confident are we in that expectation? We can use the handy `predict()` function here, which works with almost any type of model object. Check out `predict.lm` using ?, as that is the method that `predict` invokes when we feed it an lm object.

```{r}
#fitting our model
m5 <- lm(mpg ~ disp + cyl, data = mtcars)
#looking at our model
summary(m5)

#what are we doing here? Also, what does with allow us to do here?
x_vec_n <- with(mtcars, seq(min(disp), max(disp), length.out = 500))
#we've created a 500 length vector starting from the min value of disp
#through the max value of disp

#why am I choosing these values? What are other options?
new_dat_n_1 <- with(mtcars, data.frame(x_vec_n, 4)) 
new_dat_n_2 <- with(mtcars, data.frame(x_vec_n, 6))
new_dat_n_3 <- with(mtcars, data.frame(x_vec_n, 8))

#what are we doing here?
colnames(new_dat_n_1) <- c("disp", "cyl")
colnames(new_dat_n_2) <- c("disp", "cyl")
colnames(new_dat_n_3) <- c("disp", "cyl")
#we just created three data frames - all with our x_vec_n values for "disp"
#and three values we have set for "cyl"

#this is very important for predict! See note below.

#getting our predicted values 
fit_n_1 <- predict(m5, newdata = new_dat_n_1, interval = "confidence")

fit_n_2 <- predict(m5, newdata=new_dat_n_2, interval="confidence")

fit_n_3 <- predict(m5, newdata=new_dat_n_3, interval="confidence")

#ggplot is rather particular about the format of the data you will be using
plot_dat_n <- data.frame(fit_n_1, fit_n_2, fit_n_3, new_dat_n_1)
#note we never call `cyl` on its own below so the iteration of `new_dat_n_x` we include
#here doesn't matter a ton, it just gives us all values of disp

#this looks like a monster plot, but there are just a lot of layers. A lot of 
#ggplot plot code will end up looking like this
ggplot(data=plot_dat_n) + 
  labs(x = "Displacement of Vehicle", y = "Miles Per Gallon")+
  geom_line(aes(x = disp, y = fit_n_1[, 1], color = "4 cylinders"),linetype = "solid") +
  geom_line(aes(x = disp, y = fit_n_2[, 1], color = "6 cylinders"),linetype = "solid") +
  geom_line(aes(x = disp, y = fit_n_3[, 1], color = "8 cylinders"), linetype = "solid") +
  #what are we doing with the ribbons? What does alpha do? Play around with it.
  geom_ribbon(aes(x = disp, ymin = fit_n_1[, 2], ymax = fit_n_1[, 3]), alpha = .3, fill = "red") +
  geom_ribbon(aes(x = disp, ymin = fit_n_2[, 2], ymax = fit_n_2[, 3]), alpha = .3, fill = "green") +
  geom_ribbon(aes(x = disp, ymin = fit_n_3[, 2], ymax = fit_n_3[, 3]), alpha = .3, fill = "blue") +
  theme(legend.title = element_text(size = 14),
        legend.position = "bottom",
        panel.grid.minor.x = element_blank(), #what does this do?
        panel.grid.minor.y = element_blank()) +
  scale_color_discrete(name="Number of Cylinders") + #what does this seem to do? what happens if you take it out?
  theme_bw()
```

+ Note that when using predict, it is very important that our new observations have variables that bear the same name as in the formula. Otherwise `predict` won't know what to do! Try out the following:
```{r, eval = FALSE}
#recreating new_dat_n_1
new_dat_n_1_a <- with(mtcars, data.frame(x_vec_n, 4))

#but this time we don't label the variables
predict(m5, newdata = new_dat_n_1_a, interval = "confidence")

#and this time we label them incorrectly
colnames(new_dat_n_1_a) <- c("cisp", "dyl")
predict(m5, newdata = new_dat_n_1_a, interval = "confidence")

#and this time we flip the variables
new_dat_n_1_a <- with(mtcars, data.frame(4, x_vec_n))
colnames(new_dat_n_1_a) <- c("disp", "cyl")
predict(m5, newdata = new_dat_n_1_a, interval = "confidence")
#this time it works, but our predictions are clearly incorrect (compare them to
#fit_n_1 from above)

#this works!
new_dat_n_1_a <- with(mtcars, data.frame(4, x_vec_n))
colnames(new_dat_n_1_a) <- c("cyl", "disp")
fit_n_1_a <- predict(m5, newdata = new_dat_n_1_a, interval = "confidence")

all(fit_n_1 == fit_n_1_a) #what does all() do here? #check in a vector whether all the values are true or not.

```

# Part 3 Exercise: LM practice

+ Using the palmerpenguines dataset, create a linear model model using multiple predictors of your choice and then display the results in a publication quality table (using stargazer).
```{r, results='hide'}
library(palmerpenguins)
library(stargazer)

#Running model
bill_model <- lm(bill_length_mm ~ species+sex+body_mass_g, penguins)

summary(bill_model)

#Putting results in table (latex version)
stargazer(bill_model,title="Effect of Predictors on Bill Length",digits=3,star.cutoffs = c(0.1,0.05,0.01),
          label="bill_model",
          covariate.labels = c("\\textbf{Species} (reference= Adélie) &   \\\\
                      \\hspace{1cm}Chinstrap",
                      "\\hspace{1cm}Gentoo",
                    "\\textbf{Sex} (reference= Female) &   \\\\
                      \\hspace{1cm}Male",
                      "\\textbf{Body mass (grams)}"),
          notes="\\parbox[t]{12cm}{Coefficients reported from an OLS regression model. Significance codes:*$<$ 0.10,**p$<$ 0.05, ***p$<$ 0.01, two-tailed tests.}",
          notes.append=FALSE,
          style="apsr",
          dep.var.caption = c(""),
          dep.var.labels.include = FALSE,
          table.layout ="-c-!t-s-n",
          header=FALSE,
          notes.align= "l",
          column.sep.width = "1pt",
          font.size = "scriptsize",
          no.space = TRUE)
```

```{r, results='asis'}
#Putting results in table (html version)
stargazer(bill_model,title="Effect of Predictors on Bill Length",digits=3,star.cutoffs = c(0.1,0.05,0.01),
          label="bill_model",
          covariate.labels = c("\\textbf{Species} Chinstrap",
                      "\\textbf{Species} Gentoo",
                    "\\textbf{Male} (reference= Female)",
                      "\\textbf{Body mass (grams)}"),
          notes="\\parbox[t]{12cm}{Coefficients reported from an OLS regression model. Significance codes:*$<$ 0.10,**p$<$ 0.05, ***p$<$ 0.01, two-tailed tests. Species reference= Adélie}",
          notes.append=FALSE,
          style="apsr",
          dep.var.caption = c(""),
          dep.var.labels.include = FALSE,
          table.layout ="-c-!t-s-n",
          header=FALSE,
          notes.align= "l",
          column.sep.width = "1pt",
          font.size = "scriptsize",
          no.space = TRUE,
          type="html")

```

```